### 2018-07-05 paper/6-engineering.tex
Polish S6.5-6.6

To legitimise shallow builds, we need to relax the correctness
Definition~\ref{def-correct} as follows. Let the \hs{shallow} store correspond
to the result of a shallow build. Then \hs{shallow} is correct, if \emph{there
exists} a \hs{result} which satisfies all requirements of the
Definition~\ref{def-correct}, \emph{such that} \hs{shallow} agrees with the
\hs{result} on all the input keys \hs{k}~$\in$~$I$:
\[
\hs{getValue}~\hs{k}~\hs{shallow}~\hs{==}~\hs{getValue}~\hs{k}~\hs{result}
\]
and on the target \hs{key}:
\[
\hs{getValue}~\hs{key}~\hs{shallow}~\hs{==}~\hs{getValue}~\hs{key}~\hs{result}.
\]

@@

\subsection{Tracking and self-tracking}\label{sec-tracking-aspects}

@@

\begin{minted}[xleftmargin=10pt]{text}
A1 = 20      B1 = A1 + A2
A2 = 10
\end{minted}

@@

\begin{minted}{haskell}
sprsh4 "B1" = Just $ Task $ \fetch -> do
    formula <- fetch "B1-formula"
    evalFormula fetch formula
\end{minted}

% It is not possible to
% change dependencies based on the formula in a \hs{Tasks}~\hs{Applicative}, as
% would be required, so instead the formula can be captured as a dependency
% (but its value not used) and \emph{also} baked directly into the \hs{sprsh4}
% function. \simon{Don't understand.}
% \simon{I think you mean that you can find dependencies with executing the function,
%   which seems quite difference from ``backing it into'' the Haskell code.}

\noindent
The build systems that have precise self-tracking are all ones which use a
\emph{non-embedded domain specific language} to describe computations.
Those which make use of a full programming language, e.g. \Shake, are faced with
the challenge of implementing equality on arbitrary task functions. For such
build systems the incredibly pessimistic assumption of saying that any change to
the build system potentially changes any build rule can often be used~--~the
classic example being a makefile depending on itself.

@@

Some computations are best described not by a chain of acyclic dependencies,
but by a loop~--~for example \LaTeX~requires repeated rebuilding until it
reaches a fixed point~--~something that can be directly expressed in build
systems, such as \Pluto~\cite{erdweg2015pluto}. Another example of cyclic
computations is \Excel, where a cell can depend on itself, for example:

\begin{minted}[xleftmargin=10pt]{text}
A1 = A1 + 1
\end{minted}

In such cases \Excel will normally not execute anything, but if the
``Iterative Calculations'' feature is enabled will execute the formula for a
specified maximum number $N$ of times per calculation (where $N$ is a
setting that defaults to 100).

For examples like \LaTeX~we consider the proper encoding to not be with circular
tasks, but with a series of iterative steps, as described
by~\citet{shake-fixed-point}. It is important that the number of executions is
bounded, otherwise the build system may not terminate (a legitimate concern
with \LaTeX, which can be put into a situation where it is bistable or diverging
over multiple executions).

The examples in \Excel tend to encode either mutable state, or recurrence
relations. The former is only required because \Excel inherently lacks the
ability to write mutable state, and the latter is probably better solved using
explicit recurrence formulae.

>>>>

To legitimise shallow builds, we need to relax the correctness
Definition~\ref{def-correct} as follows. Let the \hs{shallow} store correspond
to the result of a shallow build. Then \hs{shallow} is correct, if \emph{there
exists} a \hs{result} which satisfies all requirements of the
Definition~\ref{def-correct}, \emph{such that} \hs{shallow} agrees with the
\hs{result} on all the input keys \hs{k}~$\in$~$I$:
\[
\hs{getValue}~\hs{k}~\hs{shallow}~\hs{==}~\hs{getValue}~\hs{k}~\hs{result}\text{,}
\]
and on the target \hs{key}:
\[
\hs{getValue}~\hs{key}~\hs{shallow}~\hs{==}~\hs{getValue}~\hs{key}~\hs{result}\text{.}
\]

@@

\subsection{Self-tracking}\label{sec-tracking-aspects}

@@

\vspace{0.5mm}
\begin{minted}[xleftmargin=10pt]{text}
A1 = 20      B1 = A1 + A2
A2 = 10
\end{minted}
\vspace{0.5mm}

@@

\vspace{0.5mm}
\begin{minted}[xleftmargin=10pt]{haskell}
sprsh4 "B1" = Just $ Task $ \fetch -> do
    formula <- fetch "B1-formula"
    evalFormula fetch formula
\end{minted}
\vspace{0.5mm}

\noindent
The build systems that have precise self-tracking are all ones which use a
\emph{non-embedded domain specific language} to describe build tasks. Those
which use a full programming language, e.g. \Shake, are faced with the challenge
of implementing equality on arbitrary task functions. For such build systems the
pessimistic assumption of saying that any change to the build system potentially
changes any build rule can often be used~--~the classic example being a makefile
depending on itself.

@@

Some computations are best described not by a chain of acyclic dependencies,
but by a loop. For example, \LaTeX~requires repeated rebuilding until it
reaches a fixed point, which can be directly expressed in build systems, such as
\Pluto~\cite{erdweg2015pluto}. Another example is \Excel, where a cell can
depend on itself, for example: \cmd{A1 = A1 + 1}. In such cases \Excel will
normally not execute anything, but if the ``Iterative Calculations'' feature is
enabled will execute the formula for a specified maximum number $N$ of times per
calculation (where $N$ is a setting that defaults to 100).

For examples like \LaTeX~we consider the proper encoding to not be with circular
tasks, but with a series of iterative steps, as described
by~\citet{shake-fixed-point}. It is important that the number of executions is
bounded, otherwise the build system may not terminate (a legitimate concern
with \LaTeX, which can be put into a situation where it is bistable or diverging
over multiple executions). The examples in \Excel tend to encode either mutable
state, or recurrence relations. The former is only required because \Excel
inherently lacks the ability to write mutable state, and the latter is probably
better solved using explicit recurrence formulae.
